---
title: "TimeSeries_Forecast"
author: "RJ"
date: "8/16/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Time Series - Forecast
This document summarizes the learning process on Time Series analysis. The original source of the learning material is from [Datacamp](https://www.datacamp.com/projects/515).


### Preparation
Load the dataset and the required packages
```{r}
library(tidyverse)
library(lubridate)
library(forecast)
library(Hmisc)
library(magrittr)

ptt_price <- read.csv(choose.files())
head(ptt_price)

```


```{r}
# Show unique value in each column
summarise_all(ptt_price, n_distinct)

# Select column with more than one unique value
ptt_price_selected <- ptt_price %>% 
  select_if(summarise_all(., n_distinct) > 1) %>% 
  rename(region_id = adm1_id,
         region = adm1_name,
         market_id = mkt_id,
         market = mkt_name,
         month = mp_month,
         year = mp_year,
         price = mp_price) %>%
  mutate(date = ymd(paste(year, month, "01"))) %>% 
  select(region_id, market, date, price)

describe(ptt_price_selected)
```

```{r}

location <- list.dirs(choose.dir())
allfile <- list.files(path = location, pattern = c(".csv"))

datalist <- list()
for (x in 1:length(allfile)) {
  # Read dataset
  datalist[[x]] <- read_csv(paste(location, allfile[x], sep = "\\"))
}

allfood <- reduce(datalist, bind_rows) %>% 
              select_if(summarise_all(., n_distinct) > 1) %>% 
              rename(region_id = adm1_id,
                     region = adm1_name,
                     market_id = mkt_id,
                     market = mkt_name, 
                     food = cm_name,
                     month = mp_month,
                     year = mp_year,
                     price = mp_price) %>%
              mutate(date = ymd(paste(year, month, "01"))) %>% 
              select(region_id, market, date, food, price)

describe(allfood)

```


### Exploration
This 

```{r}
allfood %>% 
  filter(food == "Potatoes (Irish)") %>% 
  group_by(year(date), month(date)) %>% 
  mutate(average_price = mean(price)) %>% 
  ggplot() +
    geom_line(aes(x = date,
                  y = price,
                  group = market),
              alpha = 0.1) +
    geom_line(aes(x = date,
                  y = average_price,
                  color = "red")) +
    theme_bw() +
    labs(title = "Potato Price Trend")

```


### Forecasting
Forecasting preparation. As mentioned in the source, best way to conduct forecasting is by using a `ts` object. Further research should be done to ease this process.
```{r}
# Make list
food_price_list <- allfood %>% 
  group_by(food, date) %>% 
  summarise(price = mean(price)) %>% 
  group_split(food, .keep = F) %>% 
  setNames(unique(allfood$food))

# Create ts object for forecasting
food_ts <- food_price_list$`Potatoes (Irish)` %$% 
              ts(price, 
                 start = c(year(min(date)), month(min(date))), 
                 end = c(year(max(date)), month(max(date))),
                 frequency = 12) 

food_ts
```


```{r}
food_forcast <- forecast(object = food_ts)
autoplot(food_forcast) +
  theme_bw()
```

##### *Simplify by making a function*
Using the food_price_list
```{r}

plot_predict <- function(dataset, foodLabel = "Food") {
dataset %$% 
  ts(price, 
  start = c(year(min(date)), month(min(date))), 
  end = c(year(max(date)), month(max(date))),
  frequency = 12) %>% 
  forecast() %>% 
  autoplot() +
    theme_bw() +
    labs(y = foodLabel)
}

plot_predict(dataset = food_price_list$Cassava, foodLabel = "Cassava")

```

