---
title: "RideSharingService"
author: "RJ"
date: "8/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
This machine learning exercise is documented by following a project done by AlverniaEP. Look to her [project](https://rpubs.com/aepoetry/beam_me_up) for the original verion.

The project is related to a ride sharing service in Turkey. The objective is to forecast the driver demand for the next 5 days.
There are 3 forecasting method used:
- `est`
- `stlm`
- `tbats`
For knowledge improvement, best to read [this book](https://books.google.co.id/books?id=_bBhDwAAQBAJ&hl=id&source=gbs_book_other_versions) released by the package `forecast` author.


## Preparation
Call the packages and the required dataset. The provided data has features as follow:
- `timeStamp`: Order time
- `driverID`: Driver ID
- `riderID`: Rider ID
- `orderStatus`: 
- `confirmedTimeSec`: 
- `srcGeohash`: Geographical location group(?)
- `srcLong`: Source location longitude
- `scrLat`: Source location latitude
- `destLong`: Destination location longitude
- `destLat`: Destination location latitude


```{r}
library(tidyverse)
library(lubridate)
library(magrittr)
library(forecast)
library(ggthemes)
library(plotly)


rideSharing <- read_csv(choose.files())
head(rideSharing)
```



## Data Exploration
### Understanding dataset

- `srcGeohash` and `orderStatus`
By looking to the grouping summary. It looks like that most of the order is centralized in **s** `srcGeohash` area. Other area most likely has nodrivers or canceled.

```{r}
rideSharing %>% 
  group_by(srcGeohash, orderStatus) %>% 
  summarise(total_status = n())

```

- `orderStatus`
As mentioned by the first researcher, the orderStatus feature has "cancelled" and duplicated "nodriver". BY looking to the method used to filter out the "nodriver",  assumption taken on the method would be:
- Rider only can order one at a time, especially if the order time is under the same hour. As grouping is under the same hour 

```{r}
orderData <- rideSharing %>% 
  filter(orderStatus != "cancelled") %>% 
  mutate(Date = timeStamp %>% format("%Y-%m-%d"),
         Hour = timeStamp %>% hour()) %>%
  # arrange and group by : riders, date, and hour
  arrange(riderID, Date, Hour) %>%
  group_by(riderID, Date, Hour) %>%
  mutate(prevOrder = lag(orderStatus, default = "NA"),
         dupOrder = ifelse(orderStatus == prevOrder, "Duplicates", "Not Duplicates"),
         orderStatus = ifelse(orderStatus == "nodrivers" & dupOrder == "Duplicates", NA, orderStatus)) %>%
  # drop NA in orderStatus
  filter(!is.na(orderStatus)) %>%
  # We're left with "nodrivers" and "confirmed", we still have to throw the "nodrivers" to make it 1 legitimate order
  mutate(nextOrder = lead(orderStatus, default = "NA"),
         orderStatus = ifelse(orderStatus == "nodrivers" & nextOrder == "confirmed", NA, orderStatus)) %>%
  filter(!is.na(orderStatus)) %>%
  # stop the grouping
  ungroup() %>% 
  # Get the requiered data for order forecasting
  select(timeStamp, Date, Hour, riderID, orderStatus)
head(orderData)
```


## Forecast
### Data preparation

```{r}
demandData <- orderData %>% 
                group_by(Date, Hour) %>% 
                summarise(Demand = n()) %>% 
                ungroup() %>% 
                complete(Date, Hour) %>% 
                mutate(Demand = ifelse(is.na(Demand), 0, Demand)) %>% 
                mutate(Date = as.Date(Date)) %>% 
                mutate(timeStamp = as.POSIXct(paste(Date, Hour), format = "%Y-%m-%d %H"))
head(demandData)

```

```{r}

demandData_ts <- ts(demandData$Demand, start = min(demandData$Date), frequency = 24)

demandData_ts %>%
  head(24 * 7 * 4) %>%
  decompose() %>% 
  autoplot() + # Forecast library
  theme_bw()

```

```{r}
demandData_tsSsn <- msts(demandData$Demand, seasonal.periods = c(24, 24 * 7))

demandData_tsSsn %>% 
  head(24 * 7 * 4) %>% 
  mstl() %>% 
  autoplot() +
  theme_bw()


```


```{r}

demandData %>% 
  mutate(Day = weekdays(Date)) %>% 
  ggplot(aes(x = Hour,
             y = Demand,
             fill = as.factor(Day))) +
  geom_bar(stat = "identity",
           position = position_dodge()) +
  theme_bw() +
  labs(fill = "",
       title = "Total Demand at Hour of the Day")

```


```{r}
constant <- 1

inputForecast <- head(demandData_ts + constant, length(demandData_ts) - 24*6)
inputForecast_Sn <- head(demandData_tsSsn + constant, length(demandData_tsSsn) - 24*6)
y_test <- tail(demandData_ts, 24*6)

# Transform constant
removeConstant <- function(forecastResult, constant = 1){
  forecastResult[["mean"]] <- forecastResult[["mean"]] - constant
  forecastResult[["upper"]] <- forecastResult[["upper"]] - constant
  forecastResult[["lower"]] <- forecastResult[["lower"]] - constant
  return(forecastResult)
}

```


```{r}

tsForecast <- inputForecast %>% 
  ets(lambda = 0) %>% 
  forecast(h = 24*6) %>%   
  removeConstant()

mstsForecast <- inputForecast_Sn %>% 
  stlm(lambda = 0) %>% 
  forecast(h = 24*6) %>% 
  removeConstant()

tbatsForecast <- inputForecast_Sn %>% 
  log() %>% 
  tbats(use.box.cox = FALSE)

tbatsForecast$lambda <- 0

tbatsForecast %<>%
  forecast(h = 24*6) %>% 
  removeConstant()

rbind(accuracy(round(as.vector(tsForecast$mean)), y_test),
      accuracy(round(as.vector(mstsForecast$mean)), y_test),
      accuracy(round(as.vector(tbatsForecast$mean)), y_test)) %>% 
  as_tibble(rownames = "Model") %>% 
  mutate(Model = c("ets", "stlm", "tbats"))
  
```


```{r}
forecastCompare <- data.frame(Timestamp = tail(demandData$timeStamp, 24*6),
           Actual = y_test,
           tsModel = as.vector(tsForecast$mean),
           stmlModel = as.vector(mstsForecast$mean),
           tbatsModel = as.vector(tbatsForecast$mean))

ggplotly(
  forecastCompare %>% 
  gather(key, value, Actual, tsModel, stmlModel, tbatsModel) %>% 
  ggplot(aes(x = Timestamp,
             y = value,
             color = key)) +
  geom_line() +
  theme_bw()
)
```


```{r}
forecastCompare %>% 
  mutate(tsError = Actual - tsModel,
         stmlError = Actual - stmlModel,
         tbatsError = Actual - tbatsModel) %>% 
  select(Timestamp, matches("Error")) %>% 
  gather(key, value, matches("Error")) %>% 
  ggplot(aes(x = Timestamp, 
             y = value,
             color = key,
             shape = key)) +
  geom_point(alpha = 0.4, size = 3) +
  theme_bw() 
  
ggplotly(forecastCompare %>% 
  mutate(tsError = Actual - tsModel,
         stmlError = Actual - stmlModel,
         tbatsError = Actual - tbatsModel) %>% 
  select(Timestamp, matches("Error")) %>% 
  gather(key, value, matches("Error")) %>% 
  ggplot(aes(x = Timestamp, 
             y = value,
             color = key,
             shape = key)) +
  geom_point(alpha = 0.4, size = 3) +
  theme_bw() 
)
```




## Findings and next step
- At this point stml method produce the best forecasting performance with the lowest MAE error. 
- Simplified forecasting function
```{r}

frcst <- function(forcastFeature = FALSE, 
                  timeFeature = TRUE, 
                  fun, 
                  lambda = 0,
                  h = 10, 
                  freq = 24, 
                  season = 7, 
                  addDependentVar = 1){
  
  # transform data into timeseries forecasting format
  if (deparse(substitute(fun)) == "ets") {
    input <- ts(forcastFeature, start = min(timeFeature), frequency = freq) + addDependentVar
  } else if (deparse(substitute(fun)) == "stlm") {
    input <- msts(forcastFeature, seasonal.periods = c(freq, freq * season)) + addDependentVar
  }

  # Function to add constant avoid Infinite log transformation value
  removeConstant <- function(forecastResult, constant = addDependentVar){
    forecastResult[["mean"]] <- forecastResult[["mean"]] - constant
    forecastResult[["upper"]] <- forecastResult[["upper"]] - constant
    forecastResult[["lower"]] <- forecastResult[["lower"]] - constant
    return(forecastResult)
  }
  
  # Forecasting
  forecasting <- input %>%
    fun(lambda = lambda) %>%
    forecast(h = h) %>%
    removeConstant()
  
  return(forecasting)
}
```

- Learn from the package makers. Book is available [here](https://books.google.co.id/books?id=_bBhDwAAQBAJ&hl=id&source=gbs_book_other_versions). 
